---
title: "MATH2406 TP4, 2020"
author: "Jayarajkumar Ayyappan and S3837952"
subtitle: Assignment 1 Applied analytics
output: 
  html_notebook: default
---

## Setup

Install and load the necessary packages to reproduce the report here:

```{r, echo = TRUE, warnings = FALSE}

# Already installed the below required packages in my Rstudio as part data wrangling course.
#install.packages("dplR")
#install.packages("magrittr")
#install.packages("readr")
#install.packages("xlsx")
#install.packages("tidyr")
#install.packages("lubridate")
#install.packages("Hmisc")
#install.packages("ggplot2")
#install.packages("MVN")
#install.packages("MASS")
#install.packages("infotheo")
#install.packages("outliers")
#install.packages("mlr")
#install.packages("knitr")
#install.packages("ggplot2")
```

# question 1 download the given file"pop_dataset_0002.txt" and moved to to workspace ,otherwise we need to give the complete path.

```{r}
library(readr)
df1 <- read_csv("pop_dataset_0002.txt")
View(df1)
```

```{r}
dim(df1)
str(df1)
names(df1)
 head(unique(df1$region))

```

```{r}
colSums(is.na(df1))
mean(df1$age)
min(df1$age)
max(df1$age)
sd(df1$age)

```
```{r}
q2Age <- rep (df1$age , df1$population)
q2Age
print(mean(q2Age))
print(sd(q2Age))
```


```{r}
# 3 Record the following statistics on the means from each region:
library(dplyr)
df2 <- df1 %>% group_by(region) %>% summarise(Min = min(age,na.rm = TRUE),
Q1 = quantile(age,probs = .25,na.rm = TRUE),
Median = median(age, na.rm = TRUE),
Q3 = quantile(age,probs = .75,na.rm = TRUE),
Max = max(age,na.rm = TRUE),
Mean = mean(age, na.rm = TRUE),
SD = sd(age, na.rm = TRUE),
n = n())
df2
```

```{r}
dq3 <- data.frame('region'= rep((df1$region), df1$population),
                                           'age'= rep((df1$age), df1$population))
dq3
#data frame for means for each region 
rage_mean <- dq3 %>% group_by(region) %>% summarize(Mean = mean(age))
rage_mean
print(summarise(rage_mean, mean = mean(Mean)))
print(summarise(rage_mean, sd = sd(Mean)))
print(summarise(rage_mean, Min = min(Mean)))
print(summarise(rage_mean, Q1 = quantile(Mean,0.25)))
print(summarise(rage_mean, Q2 = quantile(Mean,0.5)))
print(summarise(rage_mean, Q3 = quantile(Mean,0.75)))
print(summarise(rage_mean, Max = max(Mean)))
qplot(rage_mean$Mean,
      geom="histogram",
      binwidth = 4,  
      main = "Histogram for Mean age", 
      xlab = "Age",  
      fill=I("white"), 
      col=I("violet"), 
      alpha=I(.5))
```

```{r}
df3 <- df1 %>% group_by(region) %>% summarise(Min = min(population),
Q1 = quantile(population,probs = .25),
Median = median(population),
Q3 = quantile(population,probs = .75),
Max = max(population),
Mean = mean(population),
SD = sd(population),
SUM = sum(population),
n = n())
df3

```

```{r}
#4.1 Show which region has the least people and how many it has.
#4.2 Plot the distribution of ages for the region with the least people.
df4 <- arrange(df3,(SUM))
df4
df4 %>% filter(SUM == 3)
least_population_cities <- df3 %>% filter(SUM == 3)
joined_table <- left_join(least_population_cities, df1, 
                           by = c("region" = "region"))
age_dataset<- data.frame('Age'=rep((joined_table$age), joined_table$population))
# Histogram of the ages
qplot(age_dataset$Age,
      geom="histogram",
      binwidth = 5,  
      main = "Histogram for age of regions with the least people", 
      xlab = "Age",  
      fill=I("violet"), 
      col=I("black"), 
      alpha=I(.5))
```
```{r}
#Consider the region with the largest population:
library(ggplot2)
df5 <- arrange(df3,desc(SUM))
df5
q5 <- df1 %>% filter(region =="SSC22015")
q5
ggplot(q5, aes(population)) + stat_ecdf(geom = "step")
ggplot(q5, aes(x = population)) +
  stat_ecdf(aes(color = gender,linetype = gender), 
              geom = "step", size = 1.5) +
  scale_color_manual(values = c("#00AFBB", "#E7B800"))+
  labs(y = "f(population)")
```
Provide an analysis of trends in age against region population:
6.1 Plot the ratio of old to young people using 40 years old as a cut off, i.e. young is defined as age < 40 and old as >= 40.
6.2 Comment on any trends you see in the data. What could cause such trends?

```{r}

dfq6<-data.frame('region'=rep((df1$region), df1$population),
                                           'age'=rep((df1$age), df1$population))
dfq6

yp <-dfq6 %>% filter (age <= 40) 
yp
yplr<- yp %>% group_by(region)%>% count(region)
yplr
op <-dfq6 %>% filter (age > 40)
op
oplr <- op %>% group_by(region)%>% count(region)
oplr
Totalcount <-left_join (yplr, oplr, by = c("region" = "region"))
# we could see old people and young people count for each region side by side 
Totalcount
Totalcount$ratio <-Totalcount$n.y/Totalcount$n.x
Totalcount
Addageratio <-left_join(df3, Totalcount, by = c("region" = "region"))
Addageratio<- na.omit(Addageratio)
Addageratio
ggplot(Addageratio, aes(x = SUM, y = ratio)) +geom_point() +labs(
    title = "Plot of old to young people ratio against population",
    x = "Population",
    y = "Old to young people ratio")

```
# when the population increases old age to young age ratio decrease it indicates that, Young crowds are mostly in the city,where the population is more. I think opportunities are more in the cities and young people are migrating towards city or want to live in big cities.

# Provide an analysis of trends in gender ratio against region population:
```{r}
# prepare data for gender calculation
dq7<-data.frame('region'=rep((df1$region), df1$population),
                        'gender'=rep((df1$gender), df1$population))
dq7
males<-dq7 %>%filter(gender=='M')
males
females<-dq7 %>%filter(gender=='F')
females
mcountbyreg<-males %>%group_by(region)%>%count(region)
mcountbyreg
fcountbyreg<-females %>%group_by(region)%>%count(region)
fcountbyreg
totalgendcount <-left_join(mcountbyreg, fcountbyreg, by = c("region" = "region"))
totalgendcount<- na.omit(totalgendcount)
totalgendcount
totalgendcount$ratio<-totalgendcount$n.x/totalgendcount$n.y
genderratio <-left_join(df3,totalgendcount, by = c("region" = "region"))
genderratio<- na.omit(genderratio)
genderratio
ggplot(genderratio, aes(x = SUM, y = ratio)) +
  geom_point() +labs(
    title = "Plot of gender ratio (male to female) against region population",
    x = "Population",
    y = "Male to Female ratio")

```
# Male to femaie ratio is equal to 1 accross population and we can see interm young babies more number of males compare to females. So in the small towns or suburbs male population will be slightly higher.

```{r}
#population by gender
df8 <- df1 %>% group_by(gender) %>% summarise(Min = min(population),
SUM = sum(population),
n = n())
df8
df8byage <-df1 %>% group_by(age) %>% summarise(SUM =sum(population))
df8adesc<- arrange(df8byage,desc(SUM))
head(df8adesc)
# we will go with female gender and age group of 28 to 31 for our products as female population and togal population is more. 
print(head(fcountbyreg %>% arrange(desc(n))))
agefilter <- df1 %>% filter(28<= age & age <= 31 )
TPby_region_age<-agefilter %>% group_by(region, age) %>% summarize(SUM = sum(population))
df8b <- arrange(TPby_region_age,desc(SUM))
head(df8b)
```

```{r}
# based on the above we have selected region ssc22015
age<- df1 %>% filter(region=='SSC22015')
data_for_age<-data.frame('Region'=rep((age$region), age$population),'Age'=rep((age$age), age$population))
n<-1
xbar<-rep(NA,n)

for (i in 1: n){
  
  mysample<-sample(data_for_age$Age, size = 100)
  xbar[i]<-mean(mysample)
}

mean(xbar)
# 9.2 Add another sample of 100 ages drawn from the same sample.
n<-2
ybar<-rep(NA,n)
for (i in 1: n){
  mysample<-sample(data_for_age$Age, size = 100)
  ybar[i]<-mean(mysample)
}
mean(ybar)

sd(ybar)
hist(ybar)
lines(density(ybar), col="violet", lwd=2)
#Repeat this process for N = 1 (the original sample), 10, 100 and 1000 samples.
n<-10
zbar<-rep(NA,n)
for (i in 1: n){
  
  mysample<-sample(data_for_age$Age, size = 100)
  zbar[i]<-mean(mysample)
}
mean(zbar)
sd(zbar)
hist(zbar)
lines(density(zbar), col="violet", lwd=2)

n<-100
z1bar<-rep(NA,n)
for (i in 1: n){
  
  mysample<-sample(data_for_age$Age, size = 100)
  z1bar[i]<-mean(mysample)
}
mean(z1bar)
sd(z1bar)
hist(z1bar)
lines(density(z1bar), col="violet", lwd=2)

n<-1000
z2bar<-rep(NA,n)
for (i in 1: n){
  
  mysample<-sample(data_for_age$Age, size = 100)
  z2bar[i]<-mean(mysample)
}
mean(z2bar)
sd(z2bar)
hist(z2bar)
lines(density(z2bar), col="violet", lwd=2)
```
<br>
<br>
